---
title: "Assignment_4"
author: "Sikun_Ma"
date: '2022-07-16'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)
library(readr) 
library(dplyr) 
library(ggplot2) 
library(gridExtra) 
library(tictoc) 
library(caret)
library(psych)
library("rpart")
library("e1071")
library("randomForest")
set.seed(1)
```

## **Importing, Combining, Data Overview**

# Read data

```{r}
train_titanic <- read_csv("/Users/apple/Desktop/A04/titanic_train.csv") %>%
  mutate(Test_Data = 0)

test_titanic <- read_csv("/Users/apple/Desktop/A04/titanic_test.csv") %>%
  mutate(Test_Data = 1)

titanic_full <- bind_rows(train_titanic, test_titanic) 

glimpse(titanic_full)
```

# There are 12 variables present in the titanic_full dataset that are potentially useful. I define the function missing_vars, which I can use to get an overview of what proportion of each variable is missing, and re-use it later if I need to.

```{r}
missing_vars <- function(x) {
  var <- 0
  missing <- 0
  missing_prop <- 0
  for (i in 1:length(names(x))) {
    var[i] <- names(x)[i]
    missing[i] <- sum(is.na(x[, i]))
    missing_prop[i] <- missing[i] / nrow(x)
  }
  (missing_data <- data.frame(var = var, missing = missing, missing_prop = missing_prop) %>% 
      arrange(desc(missing_prop)))
}
```

```{r}
missing_vars(titanic_full)
```

## `Cabin` & `Age` have some significant proportion of missing values, whereas `Embarked` & `Fare` are far less of a problem.

```{r}
head(train_titanic)
```

```{r}
describe(train_titanic)
```

# A function to extract features

```{r}
extractFeatures <- function(data) {
  features <- c("Pclass",
                "Age",
                "Sex",
                "Parch",
                "SibSp",
                "Fare",
                "Embarked")
  fea <- data[,features]
  fea$Age[is.na(fea$Age)] <- -1
  fea$Fare[is.na(fea$Fare)] <- median(fea$Fare, na.rm=TRUE)
  fea$Embarked[fea$Embarked==""] = "S"
  fea$Sex      <- as.factor(fea$Sex)
  fea$Embarked <- as.factor(fea$Embarked)
  #fea <- cbind(fea, fea$Age * fea$Age)
    return(fea)
}
```

```{r}
summary(extractFeatures(train_titanic))
summary(extractFeatures(test_titanic))
```

## Fitting random forest

```{r}
rf <- randomForest(extractFeatures(train_titanic), as.factor(train_titanic $ Survived), ntree = 100, importance = TRUE)
rf
```


```{r}
# create submission file
submission <- data.frame( PassengerId= test_titanic$PassengerId )  
# using model rf fit on training data to predict test data
submission$Survived <- predict( rf, extractFeatures(test_titanic) )  
# write results to CSV file
write.csv(submission, file = "1_random_forest_r_submission.csv", row.names=FALSE)
# plot importance of preditors
imp <- importance(rf, type=1 )
imp
```

# We have successfully used the model of the training dataset to predict the testing dataset. According to the importance of preditors, the main factors which affect the survivability are: Sex, Pclass, Fare, Parch, Age, Embarked and SibSp.

## Fitting classification tree

```{r}
fol= formula( as.factor(Survived) ~ Pclass + Age + Sex + Parch + SibSp + Fare + Embarked)
fit <- rpart( fol, data=train_titanic, method= "class")
```

```{r}
print(fit)
printcp(fit)
plotcp(fit)
summary(fit)
```

# Growing a tree

```{r}
plot(fit, uniform=TRUE, main="Classification Tree")       
text(fit, use.n=FALSE, all=TRUE, cex=.8 )     
```

## Conclusion

# My random forest model shows that sex, pclass and fare are the top 3 factors that could affect one's survivability.
